Simulating High-Dimensional Multivariate Data

### 2021-02-24 010-background.Rmd
begin work converting to julia

```{r ch010-processBRCA, echo=FALSE, eval=TRUE, results='none', cache=F}
brcaRDS <- paste0( "data/brca", d, ".rds" )
if ( !file.exists( brcaRDS ) | !file.exists( './data/examplePatients.txt' ) )  {
    brca0 <- TCGA2STAT::getTCGA(disease = "BRCA",
                            data.type = "RNASeq",
                            p = 6L)
    brca1 <- as.data.frame(t(brca0$dat))
    gene_median <- apply(brca1, 2, median)
    gene_median_order <- order(gene_median, decreasing = TRUE)
    brca  <- brca1[,head(gene_median_order, d)]
    writeLines( substring( row.names( brca ), first=1, last= 12 ), './data/examplePatients.txt' ) 
    saveRDS(brca, brcaRDS)
}
brca <- readRDS(file = brcaRDS)
```

>>>>

```{r ch010-processBRCA, echo=FALSE, eval=TRUE, results='none', cache=F}
brcaRDS <- paste0( "data/brca", d, ".rds" )
if ( !file.exists( brcaRDS ) | !file.exists( './data/examplePatients.txt' ) )  {
    brca0 <- TCGA2STAT::getTCGA(disease = "BRCA",
                                data.type = "RNASeq")
    brca1 <- as.data.frame(t(brca0$dat))
    gene_median <- apply(brca1, 2, median)
    gene_median_order <- order(gene_median, decreasing = TRUE)
    brca  <- brca1[, head(gene_median_order, d)]
    writeLines(substring(row.names(brca), first = 1, last = 12), './data/examplePatients.txt') 
    saveRDS(brca, brcaRDS)
}
brca <- readRDS(file = brcaRDS)
```

### 2021-02-24 030-bigsimr-package.Rmd
begin work converting to julia

@@

```{r ch030-basic-example-setup, echo=TRUE, message=FALSE, cache=F}
## devtools::install_github("SchisslerGroup/bigsimr", ref="dev-julia")
library(bigsimr)
Sys.setenv(JULIA_NUM_THREADS=parallel::detectCores())
## Either let JuliaCall find your installation of Julia or set manually
bs <- bigsimr_setup(JULIA_HOME = "/usr/local/bin/")
dist <- distributions_setup() # Julia distributions
library(tidyverse)
library(patchwork)
```

@@

Next, we configure the input marginals for later input into `bigsimr::rvec`.
The marginal distributions are specified using `R`'s special `alist` function.
This allows one to enter the distributions without evaluating anything (yet).

```{r ch030-margins-alist, echo=TRUE, cache=T}
margins <- alist(
    qnorm(mean = 77.871, sd = 9.4855),
    qlnorm(meanlog = 3.419, sdlog = 0.6967)
)
```

Notice that we use the *quantile* function for the marginals, as that is how the marginal distributions $F_i$ are specified the `bigsimr::rvec` algorithm.
This implementation strategy supports all `R` `base` probability distributions.
It allows flexible extensions using other `R` packages that adhere to conventions, such as `extraDistr`.
Further, by using `alist`, users can specify their own custom distributions (see below in [creating custom margins](#custom-margins)).

To circumvent manual entry, `mlist` presents an alternative to `alist` with on-the-fly evaluation.

```{r ch030-margins-mlist, echo=TRUE}
mlist(
  qnorm(mean = mean(df$Temp), sd = sd(df$Temp)),
  qlnorm(meanlog = mle_mean(df$Ozone), sdlog = mle_sd(df$Ozone))
)
```

*Specifying correlation*.
As mentioned, the user must decide how to describe correlation, based on the particulars of the problem.
For non-normal data and for improved simulation accuracy/scalability in our scheme, we advocate the use of Spearman correlation matrix $R_S$ and Kendall $\tau$ matrix $R_K$.
We also support approximate Pearson correlation coefficient matching, while cautioning the user to check the performance for their parametric multivariate model (see [Monte Carlo evaluations](#simulations) below for evaluation strategies and guidance).
To aid in correlation specification, and estimation in general, we provide a high-performance function `bigsimr::cor_fast` which estimates Spearman, Kendall, or Pearson correlation using the fastest methods available.
(Anyone who has tried estimating Kendall's $\tau$ using `stats::cor` can attest that the routine does not scale to even moderate dimensions).
These estimation methods are classical approaches, not designed for high-dimensional correlation estimation (see the [Conclusion and Discussion]({#discussion) sections for more on this).

```{r ch030-aq-cor, echo=TRUE}
type <- 'spearman'
(R_S <- bs$cor(as.matrix(df), bs$Spearman))
```

*Checking target correlation matrix admissibility*. First we use `bigsimr::cor_bounds` to ensure that the pairwise target correlation values are valid prior to mapping step which constructs $R_X$ for the MVN input into NORTA (see Section \@ref(algorithms)).
`cor_bounds` estimates the pairwise lower and upper correlation bounds using the Generate, Sort, and Correlate algorithm of @DH2011.

```{r ch030-cor-bounds, echo=TRUE}
bounds <- bs$cor_bounds(D)
bounds$lower
```

Since our our single estimated Spearman correlation is within the theoretical bounds, the correlation is valid as input to `bigsimr:rvec`.
But even if the 2-dimensional bounds are satified for each pair of margins, this does not guarantee the feasibility of a $d-$variate distribution [@BF17].

To provide higher dimensional feasibility/admissibility checking, we begin by mapping using `bigsimr::cor_convert` and check admissibility.

```{r ch030-cor-convert, echo=TRUE}
## Step 1. Mapping
(R_X <- cor_convert(rho = R_S, from = type, to = 'pearson'))
## Step 2. Check admissibily
is_valid_correlation <- function(rho) {
  all(diag(rho) == 1) && matrixcalc::is.positive.semi.definite(rho)
}
is_valid_correlation( R_X )
```

@@

```{r ch030-cor-bounds-pearson, echo=TRUE}
cor_bounds(margins = margins, method = 'pearson', reps = 1e6)
```

@@

```{r ch030-sim-margins, echo=TRUE}
x <- rvec(10000, R_X, margins)
df_sim <- as.data.frame(x)
colnames(df_sim) <- colnames(df)
```

@@

*Creating a custom marginal distribution*.
Because `bigsimr` uses an `alist` to store the margins, any probability distribution with a well-defined inverse CDF can be employed, including custom marginal distributions not provided in base `R`. 
Therefore, to specify a marginal distribution absent from an existing package, the user must provide a closed-form expression to form the corresponding quantile function.

Users should prefix distributions with `r` and `q` for *random* and *quantile* respectively.
Only the quantile function is necessary to simulate random vectors, but to compute the theoretical correlation bounds, one must to supply a univariate random number generator as well.

For an example, consider a custom *Pareto* distribution for use with `bigsimr`.
The Pareto CDF is

$$
F(x) = 1 - \left(\frac{x_m}{x}\right)^\alpha
$$

for scale $x_m > 0$ and shape $\alpha > 0$, with support $x \in \left[x_m, \infty\right)$.
From the CDF, we compute the inverse CDF

$$
F^{-1}(p) = \frac{x_m}{\left(1 - p\right)^{1/\alpha}}
$$

Next we define the Pareto quantile function in R:

```{r ch030-define-qpareto, echo=TRUE}
qpareto <- function(p, scale, shape) {
  scale / (1 - p)^(1/shape)
}
```

Writing random number generating function for our target marginal can be accomplished by calling the quantile function on a uniformly distributed random variable (the inverse transform method @Rizzo2007).

```{r ch030-define-rpareto, echo=TRUE}
rpareto <- function(n, scale, shape) {
  qpareto(runif(n), scale, shape)
}
```

Now with the `qpareto` and `rpareto` functions, we can use the distribution in `bigsimr` similar to the other built-in distributions.

```{r ch030-cust-dist-example, echo=TRUE, eval=FALSE}
margins <- alist(
  qnorm(mean = 3.14, sd = 0.1),
  qbeta(shape1 = 1, shape2 = 4),
  qnbinom(size = 10, prob = 0.75),
  qpareto(scale = 1.11, shape = 5.55)
)
cor_bounds(margins, "pearson")
rho <- cor_randPD(4)
x <- rvec(10, rho, margins)
```

@@

```{r ch030-rvecRslurm-with-comments, echo=FALSE, eval=FALSE}
library(rslurm)
## a single call of rvec
## had to edit rslurm/templates/submit_single_sh.txt
## source ~/.bashrc
## conda activate bigsimr
## conda env config vars list
sjob <- slurm_call(rvec, jobname = 'rvec',
                   list(n = 1e6,
                        rho = rho,
                        margins = margins,
                        type = "spearman"),
                   submit = TRUE)
```

```{r ch030-rvecRslurm, echo=TRUE, eval=FALSE}
library(rslurm)
sjob <- slurm_call(rvec, jobname = 'rvec',
                   alist(n = 1e6, 
                         rho = rho, 
                         margins = margins, 
                         type = "spearman"),
                   submit = TRUE)
```

@@

```{r ch030-rvecRslurmMap-comments, echo=FALSE, eval=FALSE}
library(rslurm)
## a single call of rvec
## had to edit rslurm/templates/submit_sh.txt
## source ~/.bashrc
## conda activate bigsimr
## conda env config vars list
set.seed(06202020)
simReps <- 100
rhoList <- replicate(n = simReps, bigsimr::cor_randPSD(d = length(margins)),
                     simplify=FALSE)
sjob <- slurm_map(x = rhoList,
                  f = rvec,
                  jobname = 'rvecMap',
                  n = 1e6,
                  margins = margins,
                  type = "spearman",
                  nodes = 4,
                  cpus_per_node = 4,
                  submit = TRUE)

```

```{r ch030-rvecRslurmMap, echo=TRUE, eval=FALSE}
library(rslurm)
simReps <- 100
rhoList <- replicate(n = simReps, bigsimr::cor_randPSD(d = length(margins)),
                     simplify=FALSE)
sjob <- slurm_map(x = rhoList,
                  f = rvec,
                  jobname = 'rvecMap',
                  n = 1e6,
                  margins = margins,
                  type = "spearman",
                  nodes = 4,
                  cpus_per_node = 4,
                  cores = 4,
                  submit = TRUE)
```

>>>>

```{r ch030-basic-example-setup, echo=TRUE, message=FALSE, cache=F}
## devtools::install_github("SchisslerGroup/bigsimr", ref="dev-julia")
library(bigsimr)
Sys.setenv(JULIA_NUM_THREADS=parallel::detectCores())
## Either let JuliaCall find your installation of Julia or set manually
# bs <- bigsimr_setup(JULIA_HOME = "/usr/local/bin/")
bs <- bigsimr_setup(JULIA_HOME = "/home/alex/julia-1.5.3/bin/", pkg_check = FALSE)
dist <- distributions_setup() # Julia distributions
library(tidyverse)
library(patchwork)
```

@@

@@

Next, we configure the input marginals for later input into `rvec`.
The marginal distributions are specified using `Julias`'s `Distributions.jl` package and stored in an `R` vector.

```{r ch030-margins-alist, echo=TRUE, cache=T}
margins <- c(
    dist$Normal(mean(df$Temp), sd(df$Temp)),
    dist$LogNormal(mle_mean(df$Ozone), mle_sd(df$Ozone))
)
```

*Specifying correlation*.
As mentioned, the user must decide how to describe correlation, based on the particulars of the problem.
For non-normal data and for improved simulation accuracy/scalability in our scheme, we advocate the use of Spearman's $\rho$ correlation matrix $R_S$ and Kendall's $\tau$ correlation matrix $R_K$.
We also support approximate Pearson correlation coefficient matching, while cautioning the user to check the performance for their parametric multivariate model (see [Monte Carlo evaluations](#simulations) below for evaluation strategies and guidance).
To aid in correlation specification, and estimation in general, we provide a high-performance function `cor_fast` which estimates Spearman, Kendall, or Pearson correlation of large data sets by computing pairwise correlations in parallel.
(Anyone who has tried estimating Kendall's $\tau$ using `stats::cor` can attest that the routine does not scale to even moderate dimensions).
These estimation methods are classical approaches, not designed for high-dimensional correlation estimation (see the [Conclusion and Discussion]({#discussion) sections for more on this).

```{r ch030-aq-cor, echo=TRUE}
(R_S <- bs$cor(as.matrix(df), bs$Spearman))
```

*Checking target correlation matrix admissibility*.
First we use `cor_bounds` to ensure that the pairwise target correlation values are valid prior to mapping step which constructs $R_X$ for the MVN input into NORTA (see Section \@ref(algorithms)).
`cor_bounds` estimates the pairwise lower and upper correlation bounds using the Generate, Sort, and Correlate algorithm of @DH2011.

```{r ch030-cor-bounds, echo=TRUE}
bounds <- bs$cor_bounds(margins)
bounds$lower
```

Since our single estimated Spearman correlation is within the theoretical bounds, the correlation is valid as input to `rvec`.
But even if the 2-dimensional bounds are satisfied for each pair of margins, this does not guarantee the feasibility of a $d-$variate distribution [@BF17].

To provide higher dimensional feasibility/admissibility checking, we begin by mapping using `cor_convert` and check admissibility.

```{r ch030-cor-convert, echo=TRUE}
## Step 1. Mapping
(R_X <- bs$cor_convert(R_S, bs$Spearman, bs$Pearson))
## Step 2. Check admissibily
is_valid_correlation <- function(rho) {
  all(diag(rho) == 1) && matrixcalc::is.positive.semi.definite(rho)
}
is_valid_correlation(R_X)
```

@@

```{r ch030-cor-bounds-pearson, echo=TRUE}
bs$cor_bounds(margins[1], margins[2], bs$Pearson, n_samples = 1e6)
```

@@

```{r ch030-sim-margins, echo=TRUE}
x <- bs$rvec(10000, R_X, margins)
df_sim <- as.data.frame(x)
colnames(df_sim) <- colnames(df)
```

@@

@@

```{r ch030-rvecRslurm-with-comments, echo=FALSE, eval=FALSE}
library(rslurm)
## a single call of rvec
## had to edit rslurm/templates/submit_single_sh.txt
## source ~/.bashrc
## conda activate bigsimr
## conda env config vars list
sjob <- slurm_call(bs$rvec, jobname = 'rvec',
                   list(1e6,
                        rho,
                        margins),
                   submit = TRUE)
```

```{r ch030-rvecRslurm, echo=TRUE, eval=FALSE}
library(rslurm)
sjob <- slurm_call(bs$rvec, jobname = 'rvec',
                   list(1e6, 
                        rho, 
                        margins),
                   submit = TRUE)
```

@@

```{r ch030-rvecRslurmMap-comments, echo=FALSE, eval=FALSE}
library(rslurm)
## a single call of rvec
## had to edit rslurm/templates/submit_sh.txt
## source ~/.bashrc
## conda activate bigsimr
## conda env config vars list
set.seed(06202020)
simReps <- 100
rhoList <- replicate(n = simReps, bs::cor_randPSD(length(margins)),
                     simplify=FALSE)
sjob <- slurm_map(x = rhoList,
                  f = bs$rvec,
                  jobname = 'rvecMap',
                  1e6,
                  margins,
                  nodes = 4,
                  cpus_per_node = 4,
                  submit = TRUE)
```

```{r ch030-rvecRslurmMap, echo=TRUE, eval=FALSE}
library(rslurm)
simReps <- 100
rhoList <- replicate(n = simReps, bs::cor_randPSD(length(margins)),
                     simplify=FALSE)
sjob <- slurm_map(x = rhoList,
                  f = bs$rvec,
                  jobname = 'rvecMap',
                  1e6,
                  margins,
                  nodes = 4,
                  cpus_per_node = 4,
                  cores = 4,
                  submit = TRUE)
```

### 2021-02-24 040-monte-carlo-evaluation.Rmd
begin work converting to julia

@@

```{r ch040-LoadLib040, echo = FALSE, eval = FALSE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github("SchisslerGroup/bigsimr", ref="dev-julia")
library(bigsimr)
Sys.setenv(JULIA_NUM_THREADS=parallel::detectCores())
## Either let JuliaCall find your installation of Julia or set manually
bs <- bigsimr_setup(JULIA_HOME = "/usr/local/bin/")
dist <- distributions_setup() # Julia distributions
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
## cores <- as.integer( 1 )
set.seed(10162020)
```

@@

```{r ch040-biNormal, echo = FALSE, eval = TRUE}
mu <- 0
sigma <- 1
if ( !file.exists( 'results/biNorm_sims.rds' ) )  {
    mom_norm <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(mean = m, sd = s)
    }
    mu <- 0
    sigma <- 1
    margins <- alist(
        qnorm(mean = mu, sd = sigma),
        qnorm(mean = mu, sd = sigma)
    )
    type <- c("pearson", "spearman", "kendall")
    cores <- as.integer( parallel::detectCores() - 1 )
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)
    adjustForDiscrete <- c(FALSE)
    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = type, cores = cores, n = n,
                            stringsAsFactors = FALSE,
                            adjustForDiscrete = adjustForDiscrete)
    res <- data.frame()
    dir.create("./results/biNorm_sims", showWarnings = TRUE, recursive = TRUE)
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
        tmp_bounds <- cor_bounds(margins, method = type)
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)
        ## rho = cor_seq[1]
        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0
            ## Rho <- convertCor( rho = Rho, from = type, to = 'pearson' )
            time_data <- system.time({
                ## x <- rmvn(n = n, mu = rep( mu, 2 ), sigma = Rho)
                ## use rvec to test our software
                x <- rvec(n = n,
                          rho = Rho,
                          margins = margins,
                          type = type,
                          cores = cores)
            })
            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-adj", as.character(adjustForDiscrete),
                "-dev", "1CORE",
                "-lib", "bigsimr"
            )
            ## save sims in their own subdirectory
            saveRDS(x, file = paste0("./results/biNorm_sims/", id, ".rds"))
            ## Estimate statistics
            Rho_hat <- cor_fast(x, method = type)
            rho_hat <- Rho_hat[1, 2]
            norm_args_hat <- mom_norm(x[,1])
            mu_hat <- norm_args_hat$mean
            sigma_hat <- norm_args_hat$sd
            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "norm",
                                  adjustForDiscrete = adjustForDiscrete,
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  mean = mu,
                                  sd = sigma,
                                  mean_hat = mu_hat,
                                  sd_hat = sigma_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biNorm_sims.rds")
}
```

```{r ch040-biGamma, echo = FALSE, eval = TRUE}
shape <- 10
rate <- 1
if ( !file.exists( 'results/biGamma_sims.rds' ) )  {
    mom_gamma <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(shape = m^2 / s^2, rate = m / s^2)
    }
    margins <- alist(
        qgamma(shape = shape, rate = rate),
        qgamma(shape = shape, rate = rate)
    )

    type <- c("pearson", "spearman", "kendall")
    ## cores <- c(1L)
    cores <- as.integer( parallel::detectCores() - 1 )
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)
    adjustForDiscrete <- c(FALSE)

    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = type, cores = cores, n = n,
                            stringsAsFactors = FALSE,
                            adjustForDiscrete = adjustForDiscrete)

@@

        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        adjustForDiscrete <- sim_pars$adjustForDiscrete[i]

        tmp_bounds <- cor_bounds(margins, method = type)
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

        ## rho = cor_seq[1]
        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0

            time_data <- system.time({
                x <- rvec(n = n,
                          rho = Rho,
                          margins = margins,
                          cores = cores,
                          type = type,
                          ensure_PSD = TRUE)
            })

            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-adj", as.character(adjustForDiscrete),
                "-dev", "CPU",
                "-lib", "bigsimr"
            )
            saveRDS(x, file = paste0("./results/biGamma_sims/", id, ".rds"))

            ## Estimate statistics
            Rho_hat <- cor_fast(x, method = type)
            rho_hat <- Rho_hat[1, 2]
            gamma_args_hat <- mom_gamma(x[,1])
            shape_hat <- gamma_args_hat$shape
            rate_hat <- gamma_args_hat$rate

            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "gamma",
                                  adjustForDiscrete = adjustForDiscrete,
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  shape = shape,
                                  rate = rate,
                                  shape_hat = shape_hat,
                                  rate_hat = rate_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biGamma_sims.rds")
}
```

```{r ch040-biNegBin, echo = FALSE, eval = TRUE}
size <- 4
prob <- 3e-04
if ( !file.exists( 'results/biNB_sims.rds' ) )  {
    mom_nbinom <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(size = m^2 / (s^2 - m), prob = m / s^2)
    }

    margins <- alist(
        qnbinom(size = size, prob = prob),
        qnbinom(size = size, prob = prob)
    )

    type <- c("pearson", "spearman", "kendall")
    cores <- as.integer( parallel::detectCores() - 1 )
    ## cores <- cores
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)
    adjustForDiscrete <- c(FALSE)

    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = type, cores = cores, n = n,
                            stringsAsFactors = FALSE,
                            adjustForDiscrete = adjustForDiscrete)

    dir.create("./results/biNB_sims", showWarnings = TRUE, recursive = TRUE)
    res <- data.frame()
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        adjustForDiscrete <- sim_pars$adjustForDiscrete[i]
        ## AGS. 17 Oct 2020. cor_bounds failed.
        ## I'll create a minimal working example and open an issue
        ## work around for now
        if ( type == 'pearson' ) {
            tmp_bounds <- cor_bounds(margins, method = type)
            cor_lo <- tmp_bounds$lower[1,2] + eps
            cor_hi <- tmp_bounds$upper[1,2] - eps
        } else {
            cor_lo <- -1 + eps
            cor_hi <- 1 - eps
        }
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0

            time_data <- system.time({
                x <- rvec(n = n,
                          rho = Rho,
                          margins = margins,
                          cores = cores,
                          type = type,
                          ensure_PSD = TRUE)
            })

            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-adj", as.character(adjustForDiscrete),
                "-dev", "CPU",
                "-lib", "bigsimr"
            )
            saveRDS(x, file = paste0("./results/biNB_sims/", id, ".rds"))

            ## Estimate statistics
            Rho_hat <- cor_fast(x, method = type)
            rho_hat <- Rho_hat[1, 2]
            nbinom_args_hat <- mom_nbinom(x[,1])
            size_hat <- nbinom_args_hat$size
            prob_hat <- nbinom_args_hat$prob

            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "nbinom",
                                  adjustForDiscrete = adjustForDiscrete,
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  size = size,
                                  prob = prob,
                                  size_hat = size_hat,
                                  prob_hat = prob_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biNB_sims.rds")
}
```

>>>>

```{r ch040-LoadLib040, echo = FALSE, eval = FALSE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github("SchisslerGroup/bigsimr", ref="dev-julia")
library(bigsimr)
Sys.setenv(JULIA_NUM_THREADS=parallel::detectCores())
## Either let JuliaCall find your installation of Julia or set manually
# bs <- bigsimr_setup(JULIA_HOME = "/usr/local/bin/")
bs <- bigsimr_setup(JULIA_HOME = "/home/alex/julia-1.5.3/bin/", pkg_check = FALSE)
dist <- distributions_setup() # Julia distributions
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
## cores <- as.integer( 1 )
set.seed(10162020)
```

@@

@@

```{r ch040-biNormal, echo = FALSE, eval = TRUE}
mu <- 0
sigma <- 1
if (!file.exists( 'results/biNorm_sims.rds'))  {
    mom_norm <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(mean = m, sd = s)
    }
    mu <- 0
    sigma <- 1
    margins <- c(
        dist$Normal(mu, sigma),
        dist$Normal(mu, sigma)
    )
    types <- c("Pearson", "Spearman", "Kendall")
    cores <- as.integer( parallel::detectCores() - 1 )
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)
    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = types, cores = cores, n = n, stringsAsFactors = FALSE)
    res <- data.frame()
    dir.create("./results/biNorm_sims", showWarnings = TRUE, recursive = TRUE)
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        tmp_bounds <- bs$cor_bounds(margins, bs[[type]])
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)
        ## rho = cor_seq[1]
        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0
            Rho <- bs$cor_convert(Rho, bs[[type]], bs$Pearson)
            time_data <- system.time({
                ## x <- rmvn(n = n, mu = rep( mu, 2 ), sigma = Rho)
                ## use rvec to test our software
                x <- bs$rvec(n, Rho, margins)
            })
            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-dev", "1CORE",
                "-lib", "bigsimr"
            )
            ## save sims in their own subdirectory
            saveRDS(x, file = paste0("./results/biNorm_sims/", id, ".rds"))
            ## Estimate statistics
            Rho_hat <- bs$cor_fast(x, bs[[type]])
            rho_hat <- Rho_hat[1, 2]
            norm_args_hat <- mom_norm(x[,1])
            mu_hat <- norm_args_hat$mean
            sigma_hat <- norm_args_hat$sd
            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "norm",
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  mean = mu,
                                  sd = sigma,
                                  mean_hat = mu_hat,
                                  sd_hat = sigma_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biNorm_sims.rds")
}
```

```{r ch040-biGamma, echo = FALSE, eval = TRUE}
shape <- 10
rate <- 1
if (!file.exists( 'results/biGamma_sims.rds'))  {
    mom_gamma <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(shape = m^2 / s^2, rate = m / s^2)
    }
    margins <- c(
        dist$Gamma(shape, rate),
        dist$Gamma(shape, rate)
    )

    types <- c("Pearson", "Spearman", "Kendall")
    ## cores <- c(1L)
    cores <- as.integer( parallel::detectCores() - 1 )
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)

    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = types, cores = cores, n = n,
                            stringsAsFactors = FALSE)

@@

        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]

        tmp_bounds <- bs$cor_bounds(margins, bs[[type]])
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

        ## rho = cor_seq[1]
        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0
            Rho <- bs$cor_convert(Rho, bs[[type]], bs$Pearson)

            time_data <- system.time({
                x <- bs$rvec(n, Rho, margins)
            })

            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-dev", "CPU",
                "-lib", "bigsimr"
            )
            saveRDS(x, file = paste0("./results/biGamma_sims/", id, ".rds"))

            ## Estimate statistics
            Rho_hat <- bs$cor_fast(x, bs[[type]])
            rho_hat <- Rho_hat[1, 2]
            gamma_args_hat <- mom_gamma(x[,1])
            shape_hat <- gamma_args_hat$shape
            rate_hat <- gamma_args_hat$rate

            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "gamma",
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  shape = shape,
                                  rate = rate,
                                  shape_hat = shape_hat,
                                  rate_hat = rate_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biGamma_sims.rds")
}
```

```{r ch040-biNegBin, echo = FALSE, eval = TRUE}
size <- 4
prob <- 3e-04
if (!file.exists( 'results/biNB_sims.rds'))  {
    mom_nbinom <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(size = m^2 / (s^2 - m), prob = m / s^2)
    }

    margins <- c(
        dist$NegativeBinomial(size, prob),
        dist$NegativeBinomial(size, prob)
    )

    types <- c("Pearson", "Spearman", "Kendall")
    cores <- as.integer( parallel::detectCores() - 1 )
    ## cores <- cores
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)

    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = types, cores = cores, n = n,
                            stringsAsFactors = FALSE)

    dir.create("./results/biNB_sims", showWarnings = TRUE, recursive = TRUE)
    res <- data.frame()
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]

        tmp_bounds <- bs$cor_bounds(margins, bs[[type]])
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0
            Rho <- bs$cor_convert(Rho, bs[[type]], bs$Pearson)

            time_data <- system.time({
                x <- bs$rvec(n, Rho, margins)
            })

            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-dev", "CPU",
                "-lib", "bigsimr"
            )
            saveRDS(x, file = paste0("./results/biNB_sims/", id, ".rds"))

            ## Estimate statistics
            Rho_hat <- bs$cor_fast(x, bs[[type]])
            rho_hat <- Rho_hat[1, 2]
            nbinom_args_hat <- mom_nbinom(x[,1])
            size_hat <- nbinom_args_hat$size
            prob_hat <- nbinom_args_hat$prob

            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "nbinom",
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  size = size,
                                  prob = prob,
                                  size_hat = size_hat,
                                  prob_hat = prob_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biNB_sims.rds")
}
```
